Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector

### 2018-08-08 documentation/paper.tex
Cited PICO-60 paper
Filled in data tables for supervised learning

The PICO-60 bubble chamber experiment for detection of weakly interacting massive particles (WIMPs) was completed in 2017. It provides an excellent basis on which to develop and test the efficacy of such techniques as a potential tool for event discrimination, relative to the conventional techniques used during initial analysis of the experiment.

@@

$WallCut(D)$ was applied to the low-resolution Fourier transforms during early experiments (since they were used for Acoustic Parameter) to minimize the number of variables between the neural network and a proven discriminator.

Once again, configurations with and without position input were tested. This time, position corrections were optionally included when position input was not used. The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$, $FourierMLP(PosCor(\beta_{N}))$ and $FourierMLPWithPos(\beta_{N}, \chi)$. Several different resolutions $N$ of $\beta_{N}$ were used. While there were variety of network architectures tested, most had a small number of layers (on the order of 3), applied dropout and L2 regularization, and used batch normalization once again.

@@

It became evident very quickly that, without $WallCut(D)$, a high resolution was not required to get a high accuracy relative to the ground truth data. $FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29. Increasing the number of bands, even up to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42.

@@

The higher resolution demonstrated significant power when $WallCut(D)$ was not applied, and only \\ $QualCut(MultiCut(D))$ was used. $FourierMLPWithPos(\beta_{20}, \chi)$ achieved only 92\% accuracy (2 percentage points lower than the same configuration without $WallCut(D)$). $FourierMLPWithPos(\beta_{40}, \chi)$ also managed 92\% validation accuracy. However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%. However, that epoch had a relatively high standard deviation, at 0.47.

@@

The following table compares the accuracy and mean class-wise standard deviation values of various supervised learning configurations, with the original PICO-60 neural network provided as a baseline:

\begin{tabular}{|l|l|l|}
    \hline
    Configuration & Max Accuracy & Mean Class-Wise Standard Deviation \\
    \hline
    $DeepConv(\omega)$ & 77\% & TODO \\
    \hline
    $DeepConvWithPos(\omega, \chi)$ & 91\% & TODO \\
    \hline
    Existing NN TODO & X & X \\
    \hline
    AP TODO & X & X \\
    \hline
\end{tabular}

>>>>

The PICO-60 bubble chamber experiment for detection of weakly interacting massive particles (WIMPs) \cite{pico} was completed in 2017. It provides an excellent basis on which to develop and test the efficacy of such techniques as a potential tool for event discrimination, relative to the conventional techniques used during initial analysis of the experiment.

@@

$WallCut(D)$ was applied to the low-resolution Fourier transforms during early experiments (since they were used for Acoustic Parameter) to minimize the number of variables between initial neural networks and a proven discriminator.

Once again, configurations with and without position input were tested. This time, position corrections were optionally included when position input was not used. The resulting configurations are furthermore referred to as $FourierMLP(\beta_{N})$, $FourierMLP(PosCor(\beta_{N}, \chi))$ and $FourierMLPWithPos(\beta_{N}, \chi)$. Several different resolutions $N$ of $\beta_{N}$ were used. While there were variety of network architectures tested, most had a small number of layers (on the order of 3), applied dropout and L2 regularization, and used batch normalization once again.

@@

It became evident very quickly that, without $WallCut(D)$, a high resolution was not required to get a high accuracy relative to the ground truth data. $FourierMLP(\beta_{8})$ managed an excellent 98\% validation accuracy, and also produced a mean class-wise standard deviation of 0.29. Increasing the number of bands to $FourierMLPWithPos(beta_{20}, \chi)$, did not improve the results; accuracy only reached 95\%, with an increased mean standard deviation of 0.42. 40 bands reduced the accuracy once again, to 91\%.

Very interestingly, $FourierMLP(\beta_{8})$ performed better than either $FourierMLP(PosCor(\beta{8}), \chi)$ or \\ $FourierMLPWithPos(\beta{8}, \chi)$, which produced accuracy values of 97\% and 95\%, respectively.

@@

The higher resolution demonstrated significant power when $WallCut(D)$ was not applied, and only \\ $QualCut(MultiCut(D))$ was used. $FourierMLPWithPos(\beta_{20}, \chi)$ achieved only 92\% accuracy (2 percentage points lower than the same configuration without $WallCut(D)$). However, training on $\beta_{50,001}$, which does not incorporate any banding, improved this figure to a maximum of 97\%. However, that epoch had a relatively high standard deviation, at 0.47.

@@

The following tables compare the accuracy and mean class-wise standard deviation values of various supervised learning configurations, with Acoustic Parameter and the original PICO-60 neural network provided as baselines.

With $WallCut(D)$:

\begin{tabular}{|l|l|l|}
    \hline
    Configuration & Max Accuracy & Mean Class-Wise Standard Deviation \\
    \hline
    $FourierMLP(PosCor(\beta{8}), \chi)$ & 97\% & 0.42 \\
    \hline
    $FourierMLP(\beta{8})$ & 98\% & 0.29 \\
    \hline
    $FourierMLPWithPos(\beta{8}, \chi)$ & 95\% & 0.44 \\
    \hline
    $FourierMLPWithPos(\beta{20}, \chi)$ & 94\% & 0.52 \\
    \hline
    $FourierMLPWithPos(\beta{40}, \chi)$ & 91\% & 0.54 \\
    \hline
    Original PICO-60 Neural Network & 94\% & 0.99 \\
    \hline
    Acoustic Parameter & N/A & 0.56 \\
    \hline
\end{tabular}

Without $WallCut(D)$:

\begin{tabular}{|l|l|l|}
    \hline
    Configuration & Max Accuracy & Mean Class-Wise Standard Deviation \\
    \hline
    $FourierMLPWithPos(\beta_{20}, \chi)$ & 92\% & 0.63 \\
    \hline
    $FourierMLPWithPos(\beta{50,001}, \chi)$ & 97\% & 0.47 \\
    \hline
    $DeepConv(\omega)$ & 77\% & 0.86 \\
    \hline
    $DeepConvWithPos(\omega, \chi)$ & 91\% & 0.61 \\
    \hline
    $ImageConv(\iota)$ & 85\% & 0.95 \\
    \hline
    Acoustic Parameter & N/A & 0.73 \\
    \hline
\end{tabular}
